---
title: "Project full fitting process and supplementary models sta 444"
author: "Xiaoyang Liu"
date: "Due 5/3/2018"
output:
  pdf_document: default
  html_document:
    theme: cosmo
---

```{r setup, include=FALSE}
library(dplyr)
library(ggplot2)
library(tuneR)
library(forecast)
```
Dataset description. http://www.isophonics.net/SingingVoiceDataset

Research question. Do human singing voices, in particular Beijing Opera, have a simple ARIMA model? If so, is there any difference between voices of different singers and between singers from two genders? How is the difference reflected in the model's fitted parameters? 


\textbf{Full fitting process for all 6 song segments.}

Song segment #1
Singer. Female 1. Zhang Shuo
Song type. Beijing Opera
Song name. bei4 jiu1 chan2
Emotion. Postive.

```{r}
song1 <- readWave('female1_pos_1.wav')
song1left <- song1@left
s1left <- song1left / 2^(song1@bit -1)

```
```{r}
#full duration
forecast::ggtsdisplay(s1left, points=FALSE)
```
An echo pattern is observed in ACF.Don't see periodic pattern.
Because the song duration varies from one song to another and in general 1-2 mins long of recording has too many data points. I will select the beginning 5 s duration of every song for analysis.

```{r}
#5s duration at 44.1kHz sampling rate
forecast::ggtsdisplay(s1left[20000:240500], points=FALSE)
```


```{r}
#
#f1.m1=forecast::Arima(gas,order = c(0,1,0), seasonal=list(order=c(0,0,0),period=0));
f1.m1=forecast::Arima(s1left[20000:240500],order = c(0,0,0));
f1.m1
f1.m2=forecast::Arima(s1left[20000:240500],order = c(0,1,0));
f1.m2
f1.m3=forecast::Arima(s1left[20000:240500],order = c(0,2,0));
f1.m3
f1.m4=forecast::Arima(s1left[20000:240500],order = c(0,3,0));
f1.m4
```

AIC and BIC decreases with first and second order of differencing. And they don't change with third order of differencing. So i will choose second order differencing for this song.
```{r}
forecast::ggtsdisplay(f1.m1$residuals, points=FALSE)
forecast::ggtsdisplay(f1.m2$residuals, points=FALSE)
forecast::ggtsdisplay(f1.m3$residuals, points=FALSE)
forecast::ggtsdisplay(f1.m4$residuals, points=FALSE)
```
For d=2, the PACF of residual plot have strong spikes, suggesting addition of a MA term. 
```{r}
f1.2m1=forecast::Arima(s1left[20000:240500],order = c(0,2,1));
f1.2m1
f1.2m2=forecast::Arima(s1left[20000:240500],order = c(0,2,2));
f1.2m2
```
```{r}
forecast::ggtsdisplay(f1.2m1$residuals, points=FALSE)
```
Adding first and second degree MA terms, and select first degree MA terms over second.

The remaining is to test out possibility of AR terms or seasonal terms. None is found.

```{r}
f1.2m3=forecast::Arima(s1left[20000:240500],order = c(1,2,0));
f1.2m3
f1.2m4=forecast::Arima(s1left[20000:240500],order = c(2,2,0));
f1.2m4
```


```{r}
f1.3m1=forecast::Arima(s1left[20000:240500],order = c(0,2,1),seasonal=list(order=c(1,0,0),period=25));
f1.3m1
f1.3m2=forecast::Arima(s1left[20000:240500],order = c(0,2,1), seasonal=list(order=c(0,0,1),period=25));
f1.3m2
```

```{r}
forecast::ggtsdisplay(f1.3m1$residuals, points=FALSE)
forecast::ggtsdisplay(f1.3m2$residuals, points=FALSE)
```

Brief summary. Its a (0,2,1) ARMA model.

Song segment #2
Singer. Female 1. Zhang Shuo
Song type. Beijing Opera
Song name. bei4 jiu1 chan2
Emotion. Postive.

Analyzing a different part of the same song as song segment#2.


```{r}
#5s duration at 44.1kHz sampling rate
forecast::ggtsdisplay(s1left[1920000:2140500], points=FALSE)
```

```{r}
#
#f1.m1=forecast::Arima(gas,order = c(0,1,0), seasonal=list(order=c(0,0,0),period=0));
f1s2.m1=forecast::Arima(s1left[1920000:2140500],order = c(0,0,0));
f1s2.m1
f1s2.m2=forecast::Arima(s1left[1920000:2140500],order = c(0,1,0));
f1s2.m2
f1s2.m3=forecast::Arima(s1left[1920000:2140500],order = c(0,2,0));
f1s2.m3
f1s2.m4=forecast::Arima(s1left[1920000:2140500],order = c(0,3,0));
f1s2.m4
f1s2.m5=forecast::Arima(s1left[1920000:2140500],order = c(0,4,0));
f1s2.m5
```
Differencing 2 or 3 times. Third order differencing gives only about 2% of improvment in AIC BIC over second order.

```{r}
forecast::ggtsdisplay(f1s2.m3$residuals, points=FALSE)
forecast::ggtsdisplay(f1s2.m4$residuals, points=FALSE)
```



```{r}
f1s2.2m1=forecast::Arima(s1left[1920000:2140500],order = c(0,2,1));
f1s2.2m1
f1s2.2m2=forecast::Arima(s1left[1920000:2140500],order = c(0,2,2));
f1s2.2m2
f1s2.2m3=forecast::Arima(s1left[1920000:2140500],order = c(0,3,1));
f1s2.2m3
f1s2.2m4=forecast::Arima(s1left[1920000:2140500],order = c(0,3,2));
f1s2.2m4
```
(0,2,2) model has the best AIC and BIC.

```{r}
f1s2.3m1=forecast::Arima(s1left[1920000:2140500],order = c(1,2,1));
f1s2.3m1
f1s2.3m2=forecast::Arima(s1left[1920000:2140500],order = c(1,2,2));
f1s2.3m2
f1s2.3m3=forecast::Arima(s1left[1920000:2140500],order = c(1,3,1));
f1s2.3m3
f1s2.3m4=forecast::Arima(s1left[1920000:2140500],order = c(1,3,2));
f1s2.3m4

```

No improvement seen from adding AR terms.

So decide on (0,2,2).

Song segment #3
Singer. Female 1. Zhang Shuo
Song type. Beijing Opera
Song name. zhe4 cai2 shi4
Emotion. Postive.


```{r}
f1.7 <- readWave('female1_pos_7.wav')
f1.7l <- f1.7@left
f1.7l <- f1.7l / 2^(f1.7@bit -1) #converting to doubles

```


```{r}
#full duration
forecast::ggtsdisplay(f1.7l, points=FALSE)
```

```{r}
#5 s
forecast::ggtsdisplay(f1.7l[1350000:1570500], points=FALSE)
```


```{r}
#
#f1.m1=forecast::Arima(gas,order = c(0,1,0), seasonal=list(order=c(0,0,0),period=0));
f1s3.m1=forecast::Arima(f1.7l[1350000:1570500],order = c(0,0,0));
f1s3.m1
f1s3.m2=forecast::Arima(f1.7l[1350000:1570500],order = c(0,1,0));
f1s3.m2
f1s3.m3=forecast::Arima(f1.7l[1350000:1570500],order = c(0,2,0));
f1s3.m3
f1s3.m4=forecast::Arima(f1.7l[1350000:1570500],order = c(0,3,0));
f1s3.m4
f1s3.m5=forecast::Arima(f1.7l[1350000:1570500],order = c(0,4,0));
f1s3.m5
```

Either second or third degree of differencing.

```{r}
forecast::ggtsdisplay(f1s3.m3$residuals, points=FALSE)
forecast::ggtsdisplay(f1s3.m4$residuals, points=FALSE)

```
It can probably go either AR or MA. Testing out both.
Models with d=3
```{r}
#
#f1.m1=forecast::Arima(gas,order = c(0,1,0), seasonal=list(order=c(0,0,0),period=0));
f1s3.2m1=forecast::Arima(f1.7l[1350000:1570500],order = c(0,3,1));
f1s3.2m1
f1s3.2m2=forecast::Arima(f1.7l[1350000:1570500],order = c(0,3,2));
f1s3.2m2
f1s3.2m3=forecast::Arima(f1.7l[1350000:1570500],order = c(0,3,3));
f1s3.2m3
f1s3.2m4=forecast::Arima(f1.7l[1350000:1570500],order = c(1,3,0));
f1s3.2m4
f1s3.2m5=forecast::Arima(f1.7l[1350000:1570500],order = c(2,3,0));
f1s3.2m5
```

Select (0,3,1) which has the least number of parameters and comparable AIC BIC as (0,3,2) and (0,3,3) with d=3.
Then proceed to models with d=2, realize (0,2,2) is actually a better model than (0,3,1)
```{r}
#
#f1.m1=forecast::Arima(gas,order = c(0,1,0), seasonal=list(order=c(0,0,0),period=0));
f1s3.3m1=forecast::Arima(f1.7l[1350000:1570500],order = c(0,2,1));
f1s3.3m1
f1s3.3m2=forecast::Arima(f1.7l[1350000:1570500],order = c(0,2,2));
f1s3.3m2
f1s3.3m3=forecast::Arima(f1.7l[1350000:1570500],order = c(1,2,1));
f1s3.3m3
f1s3.3m4=forecast::Arima(f1.7l[1350000:1570500],order = c(1,2,2));
f1s3.3m4
f1s3.3m5=forecast::Arima(f1.7l[1350000:1570500],order = c(1,3,1));
f1s3.3m5
```
No improvement from any AR terms.

```{r}
f1s3.4m1=forecast::Arima(f1.7l[1350000:1570500],order = c(0,2,2));
f1s3.4m1
```
So (0,2,2) model fits this song segment well.

```{r}

forecast::ggtsdisplay(f1s3.4m1$residuals, points=FALSE)
```

Song Segment #4
Singer. Female 10. Liang Sun
Song type. Beijing Opera
Song name. Mu Gui Ying Gua Shuai/Lady General Mu takes command
Emotion. Postive.

```{r}
f10.3 <- readWave('female10_pos_3.wav')
f10.3l <- f10.3@left
f10.3l <- f10.3l / 2^(f10.3@bit -1) #converting to doubles

```

```{r}
#full duration
forecast::ggtsdisplay(f10.3l, points=FALSE)
```
The first 1 million datapoints are just white noise. The singing starts at around 22 s into the recording.
```{r}
#5s duration
forecast::ggtsdisplay(f10.3l[1020000:1240500], points=FALSE)
```
```{r}
f10.m1=forecast::Arima(f10.3l[1020000:1240500],order = c(0,0,0));
f10.m1
f10.m2=forecast::Arima(f10.3l[1020000:1240500],order = c(0,1,0));
f10.m2
f10.m3=forecast::Arima(f10.3l[1020000:1240500],order = c(0,2,0));
f10.m3
f10.m4=forecast::Arima(f10.3l[1020000:1240500],order = c(0,3,0));
f10.m4
f10.m5=forecast::Arima(f10.3l[1020000:1240500],order = c(0,4,0));
f10.m5
```
Again select second order differencing. Since there is no significant improvements going from second to third order differencing.Residuals of (0,2,0) is plotted below.

```{r}
forecast::ggtsdisplay(f10.m3$residuals, points=FALSE)
forecast::ggtsdisplay(f10.m4$residuals, points=FALSE)
```

```{r}
f10.2m1=forecast::Arima(f10.3l[1020000:1240500],order = c(0,2,1));
f10.2m1
f10.2m2=forecast::Arima(f10.3l[1020000:1240500],order = c(0,3,1));
f10.2m2
f10.2m3=forecast::Arima(f10.3l[1020000:1240500],order = c(0,2,2));
f10.2m3
f10.2m4=forecast::Arima(f10.3l[1020000:1240500],order = c(0,3,2));
f10.2m4

```

Choose (0,3,1) over (0,3,2) because the improvement is less than 1%.
But (0,2,2) is also slightlt better than (0,3,1).
And parameters of (0,2,2)don't match the parameters of (0,2,2) models of segment#1,#2,#3.


```{r}
forecast::ggtsdisplay(f10.2m2$residuals, points=FALSE)
forecast::ggtsdisplay(f10.2m4$residuals, points=FALSE)
```
Song Segment#5
Singer 3 profile.
Male 12. Bidong Guo
Song type. Beijing Opera
Song name. Kong Cheng Ji/The stratagem of the empty city
Emotion. Postive.

```{r}
m12.1 <- readWave('male12_pos_1.wav')
m12.1l <- m12.1@left
m12.1l <- m12.1l / 2^(m12.1@bit -1) #converting to doubles

```

```{r}
#full duration
forecast::ggtsdisplay(m12.1l, points=FALSE)
```


```{r}
#5 s
forecast::ggtsdisplay(m12.1l[1020000:1240500], points=FALSE)
```


```{r}
m12.m1=forecast::Arima(m12.1l[661500:882000],order = c(0,0,0));
m12.m1
m12.m2=forecast::Arima(m12.1l[661500:882000],order = c(0,1,0));
m12.m2
m12.m3=forecast::Arima(m12.1l[661500:882000],order = c(0,2,0));
m12.m3
m12.m4=forecast::Arima(m12.1l[661500:882000],order = c(0,3,0));
m12.m4
m12.m5=forecast::Arima(m12.1l[661500:882000],order = c(0,4,0));
m12.m5
```

Second order differencing gives significant improvement over first order, while third order differencing gives minor improvement and fourth order shows no improvement.



Second order differencing has higher AIC/BIC than first order differencing. Selecting first order differencing.
```{r}
forecast::ggtsdisplay(m12.m3$residuals, points=FALSE)
forecast::ggtsdisplay(m12.m4$residuals, points=FALSE)
```
Observe greater spikes in PACF anding MA term of degree one while trying out AR term of degree 1 as well.
```{r}
m12.2m1=forecast::Arima(m12.1l[661500:882000],order = c(0,2,1));
m12.2m1
m12.2m2=forecast::Arima(m12.1l[661500:882000],order = c(0,3,1));
m12.2m2
m12.2m3=forecast::Arima(m12.1l[661500:882000],order = c(1,2,0));
m12.2m3
m12.2m4=forecast::Arima(m12.1l[661500:882000],order = c(1,3,0));
m12.2m4
```

Select (0,3,1) with best AIC and BIC.


```{r}
m12.3m1=forecast::Arima(m12.1l[661500:882000],order = c(0,2,2));
m12.3m1
m12.3m2=forecast::Arima(m12.1l[661500:882000],order = c(3,3,1));
m12.3m2
m12.3m3=forecast::Arima(m12.1l[661500:882000],order = c(0,3,2));
m12.3m3
m12.3m4=forecast::Arima(m12.1l[661500:882000],order = c(0,3,3));
m12.3m4
```

(0,3,2) gives minor improvement of both AIC and BIC over (0,3,1). 

But again (0,2,2) gives better result than (0,3,2). So (0,2,2) is selected eventually.

```{r}
m12.3m1
forecast::ggtsdisplay(m12.3m1$residuals, points=FALSE)

```



Song segment#6
Singer Profile. 
Male 13 Tingfeng Zhang
Song type. Beijing Opera
Song name. Si Lang Tan Mu/Yang Silang Visits His Mother
Emotion. Postive.

```{r}
m13.3 <- readWave('male13_pos_3.wav')
m13.3l <- m13.3@left
m13.3l <- m13.3l / 2^(m13.3@bit -1) #converting to doubles

```

```{r}
#full duration
forecast::ggtsdisplay(m13.3l, points=FALSE)
```



```{r}
forecast::ggtsdisplay(m13.3l[50000:270500], points=FALSE)
```

```{r}
m13.m1=forecast::Arima(m13.3l[50000:270500],order = c(0,0,0));
m13.m1
m13.m2=forecast::Arima(m13.3l[50000:270500],order = c(0,1,0));
m13.m2
m13.m3=forecast::Arima(m13.3l[50000:270500],order = c(0,2,0));
m13.m3
m13.m4=forecast::Arima(m13.3l[50000:270500],order = c(0,3,0));
m13.m4
m13.m5=forecast::Arima(m13.3l[50000:270500],order = c(0,4,0));
m13.m5
```
Select either second or third degree of differencing.
```{r}
forecast::ggtsdisplay(m13.m3$residuals, points=FALSE)
forecast::ggtsdisplay(m13.m4$residuals, points=FALSE)

```



```{r}
m13.2m1=forecast::Arima(m13.3l[50000:270500],order = c(0,2,1));
m13.2m1
m13.2m2=forecast::Arima(m13.3l[50000:270500],order = c(0,3,1));
m13.2m2
m13.2m3=forecast::Arima(m13.3l[50000:270500],order = c(0,2,2));
m13.2m3
m13.2m4=forecast::Arima(m13.3l[50000:270500],order = c(0,3,2));
m13.2m4
m13.2m5=forecast::Arima(m13.3l[50000:270500],order = c(1,2,0));
m13.2m5
m13.2m5=forecast::Arima(m13.3l[50000:270500],order = c(1,3,0));
m13.2m5
```
```{r}
forecast::ggtsdisplay(m13.2m3$residuals, points=FALSE)

```

```{r}
m13.2m6=forecast::Arima(m13.3l[50000:270500],order = c(2,2,2));
m13.2m6
m13.2m7=forecast::Arima(m13.3l[50000:270500],order = c(1,3,2));
m13.2m7
m13.2m8=forecast::Arima(m13.3l[50000:270500],order = c(3,3,2));
m13.2m8
```
Decide on (0,2,2) whose AIC and BIC values are better than (0,3,2).




